---
title: "Check-in 3"
output: pdf_document
urlcolor: blue
---

  ```{r global_options, include=FALSE}
library(knitr)
library(tidyverse)
opts_chunk$set(fig.align="center", fig.height=4, fig.width=5.5)
```

**Please submit as a pdf file on Canvas. Insert answers inside the R Markdown code chunk so that the PDF shows both the code and the output.**

**Problem 1: (7 pt)**  Inside the code chunk below, write a single command (using pipes) that assigns to the object `dog_data` the following in this order:

1. Imports `stevens_etal_2020_obed_data1.csv` from the following URL: 

`https://decisionslab.unl.edu/data/stevens_etal_2020_obed_data1.csv`

2. Includes only the follow variables in this order: `date`, `class`, `dog_sex`, `cgc_test`, all four `cort` columns (using a helper function).

3. Relabels `cgc_test` to `cgc`.

4. Excludes observations where the dog's sex is missing.

5. Creates a new column called `log_cort4` that applies the `log()` function to `cort4`.

6. Sorts the data by `dog_sex` then `cort1`.

7. Gives a `glimpse()` of the data.


```{r, message = FALSE}

```


**Problem 2: (2 pts)** Inside the code chunk below, write a single command (using pipes) that calculates the mean, standard deviation, and sample size of the `cort1` variable aggregated by dog sex and CGC status for the `dog_data` object. *Note for this and the remaining questions, do not assign your output to an object.*

```{r, message = FALSE}

```



**Problem 3: (3 pts)**  Inside the code chunk below, write a single command (using pipes) that does the following to `dog_data`:

1. Creates an `id` variable that is a sequence from 1 to the length of the data frame.

2. Makes the `id` variable the first column.

3. Reshapes the data frame to be tidy such that the four `cort` columns are turned into a column of labels called `time` and a column of values called `cort`.


```{r}

```


**Problem 4: (2 pts)**  Inside the code chunk below, write a single command (using pipes) that does the following to `dog_data`:

1. Divides up the `date` variable into `year`, `month`, and `day`.

2. Creates a new column called `session` that combines the year with the class separated by a slash `/` and keeps the `year` and `class` columns.

```{r}

```


**Problem 5: (2 pts)**  Looks like more data have come in. After defining `new_data` in the code chunk below, append these new data to the bottom of `dog_data` and sort by _reverse_ chronological order by date. 
```{r}
new_data <- tibble(date = as.Date("2021-05-18"), class = "U21", dog_sex = "Female",
  cgc = "Pass", cort1 = 0.254, cort2 = NA, cort3 = NA, cort4 = 0.188, log_cort4 = -1.671313)

```


**Problem 6: (2 pts)**  The location data for each class was saved in a different data frame. After defining `location` in the code chunk below, merge the `location` data frame with `dog_data` based on `class` and then move the `site` column after the `class` column. 

```{r}
location <- tibble(class = c("U18", "U19", "S19", "F18", "S18", "F19"), 
  site = c("south", "south", "campus", "campus", "campus", "campus"))

```

