---
title: "Visualizing distributions"
output: 
  learnr::tutorial:
    progressive: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(palmerpenguins)
gradethis::gradethis_setup()
knitr::opts_chunk$set(echo = FALSE)
```


## Introduction

This exercise covers visualizing distributions including histograms, density plots, violin plots, and boxplots.

### Set-up

Let's load `{tidyverse}` and `{palmerpenguins}`.

```{r echo = TRUE, message = FALSE}
library(tidyverse)
library(palmerpenguins)
```


## Histograms

###

We make histograms with the `geom_histogram()` geometric object.

```{r hist, echo=TRUE, warning=FALSE}
penguins %>% 
  ggplot(aes(x = body_mass_g)) +
  geom_histogram()
```

### Adjust bins

As the message mentions, the default number of bins for histograms is 30. This is not necessarily ideal, so let's see what the histogram looks like with 15 bins.

```{r hist-bin, warning=FALSE}
penguins %>%    
  ggplot(aes(x = body_mass_g)) +
  geom_histogram(bins = 15)
```

**Reproduce this figure by making a histogram with 15 bins.**

```{r exercise1, exercise = TRUE, warning = FALSE}
penguins %>% 
  ggplot(...)
```

```{r exercise1-solution}
penguins %>% 
  ggplot(aes(x = body_mass_g)) +
  geom_histogram(bins = 15)
```


### Adding color to histograms

The default colors for histograms are boring. They use a dark gray both for the bars and the borders. Let's liven it up a bit.

```{r hist-color, warning=FALSE}
penguins %>%    
  ggplot(aes(x = body_mass_g)) +
  geom_histogram(bins = 15, color = "black", fill = "steelblue")
```

**Reproduce this figure by making the bars *steelblue* and the borders *black* (and keep bins at 15).**

```{r exercise2, exercise = TRUE, warning = FALSE}
penguins %>% 
  ggplot(...)
```

```{r exercise2-solution}
penguins %>% 
  ggplot(aes(x = body_mass_g)) +
  geom_histogram(bins = 15, color = "black", fill = "steelblue")
```

## Density plots

###

Histograms can be tricky to tweak to get a good bin size. An alternative that smooths over the data more is the density plot that uses the `geom_density()` function.

```{r density, echo=TRUE, warning=FALSE}
penguins %>%    
  ggplot(aes(x = body_mass_g)) +
  geom_density()
```

### Adding color to density plots

By default,  only the density line is plotted, so you may want to add color to the area below the line. If we're adding color to an area, which argument controls this?

```{r density-fill, warning=FALSE}
penguins %>%    
  ggplot(aes(x = body_mass_g)) +
  geom_density(fill = "steelblue")
```

**Color the area below the line *steelblue*.**

```{r exercise3, exercise = TRUE, warning = FALSE}
penguins %>% 
  ggplot(...)
```

```{r exercise3-solution}
penguins %>%    
  ggplot(aes(x = body_mass_g)) +
  geom_density(fill = "steelblue")
```

### Mapping colors with multiple distributions

The plots so far have included data from all three penguins species combined. But we may want to see if the species have different distributions of body mass. How do we create density plots for each of the species?

```{r density-species, warning=FALSE}
penguins %>%    
  ggplot(aes(x = body_mass_g, fill = species)) +
  geom_density()
```

**Create this plot where each species has its own density plot of a different color.**

```{r exercise4, exercise = TRUE, warning = FALSE}
penguins %>% 
  ggplot(...)
```

```{r exercise4-solution}
penguins %>%    
  ggplot(aes(x = body_mass_g, fill = species)) +
  geom_density()
```

### Changing transparency

Though the distributions are separated for each species, there is a lot of overlap between the species, with some data hidden behind others. Let's use transparency to better visualize the distributions.

```{r density-transparency, warning=FALSE}
penguins %>%    
  ggplot(aes(x = body_mass_g, fill = species)) +
  geom_density(alpha = 0.25)
```

**Create this plot where each species has its own density plot of a different color and transparency is set to 0.25.**

```{r exercise5, exercise = TRUE, warning = FALSE}
penguins %>% 
  ggplot(...)
```

```{r exercise5-solution}
penguins %>%    
  ggplot(aes(x = body_mass_g, fill = species)) +
  geom_density(alpha = 0.25)
```


## Violin plots

### 

Histograms and density plots can represent one or a few distributions, but comparing multiple distributions can be difficult with these plots. Violin plots provide a better visualization of multiple distributions. Violin plots are basically density plots that are rotated and mirrored to provide a symmetrical distribution of the data.

Violin plots use the `geom_violin()` geometric object, and here we create one for the body mass data for each species.

```{r violin, echo=TRUE, warning=FALSE}
penguins %>%    
  ggplot(aes(x = species, y = body_mass_g)) +
  geom_violin()
```

### Properties refresher

Just as a refresher, let's adjust both global and mapped properties of the plot.

```{r violin-color, warning=FALSE}
penguins %>%    
  ggplot(aes(x = species, y = body_mass_g, fill = species, color = species)) +
  geom_violin(alpha = 0.25, show.legend = FALSE)
```

**Map both violin area and borders to species, reduce transparency to 0.25, and remove the legend.**

```{r exercise9, exercise = TRUE, warning = FALSE}
penguins %>% 
  ggplot(...)
```

```{r exercise9-solution}
penguins %>%    
  ggplot(aes(x = species, y = body_mass_g, fill = species, color = species)) +
  geom_violin(alpha = 0.25, show.legend = FALSE)
```

### Adding other stats

Often it is useful to add other statistics to the plot instead of just violin plots.

```{r violin-statsummary, warning=FALSE}
penguins %>%    
  ggplot(aes(x = species, y = body_mass_g)) +
  geom_violin(fill = "lightblue", color = "steelblue", size = 1.25) +
  stat_summary(fun.data = mean_cl_normal, color = "steelblue")
```

**Add a point for the mean and error bars for 95% confidence intervals to each violin. Make the area color *lightblue*, the borders *steelblue* and size 1.25, and points/error bars *steelblue* for all violins.**

```{r exercise6, exercise = TRUE, warning = FALSE}
penguins %>% 
  ggplot(...)
```

```{r exercise6-solution}
penguins %>%    
  ggplot(aes(x = species, y = body_mass_g)) +
  geom_violin(fill = "lightblue", color = "steelblue", size = 1.25) +
  stat_summary(fun.data = mean_cl_normal, color = "steelblue")
```


## Boxplots

###

While histograms, density plots, and violin plots show fairly fine-grained representations, sometimes you want a courser representation. Boxplot provide these by showing the median, interquartile range, and other measures of range.

```{r boxplot, echo=TRUE, warning=FALSE}
penguins %>%    
  ggplot(aes(x = species, y = body_mass_g)) +
  geom_boxplot(fill = "lightblue", color = "steelblue")
```

The default behavior of boxplots is for the bar to represent the median, the box to represent the interquartile range (25th-75th percentile of the data), the whiskers to represent 1.5 times the interquartile range, and the points represent outliers (data outside of 1.5 times the interquartile range).

### Whisker length

You can change the whiskers to be the full range by specifying a very high value in `coef`, which is the value multiplied by the interquartile range to produce the whiskers (by default, `coef = 1.5`).

```{r boxplot-coef, echo=TRUE, warning=FALSE}
penguins %>%    
  ggplot(aes(x = species, y = body_mass_g)) +
  geom_boxplot(coef = 5)
```

### Outlier properties

To change properties of the outlier data points, you can append `outlier.` to the standard properties (e.g., `outlier.color`).

```{r boxplot-outlier, warning=FALSE}
penguins %>%    
  ggplot(aes(x = species, y = body_mass_g)) +
  geom_boxplot(outlier.shape = 8, outlier.size = 3)
```

**Reproduce the outlier symbols and increase the size to 3. You may have to search for R symbols to find the symbol number.**

```{r exercise7, exercise = TRUE, warning = FALSE}
penguins %>% 
  ggplot(...)
```

```{r exercise7-solution}
penguins %>%    
  ggplot(aes(x = species, y = body_mass_g)) +
  geom_boxplot(outlier.shape = 8, outlier.size = 3)
```


### Notched boxplots

You can add notches to boxplots that represent roughly 95% confidence intervals around the median by setting `notch = TRUE`.

```{r boxplot-notched, warning=FALSE}
penguins %>%    
  ggplot(aes(x = species, y = body_mass_g)) +
  geom_boxplot(notch = TRUE)
```

**Set `notch = TRUE`.**

```{r exercise8, exercise = TRUE, warning = FALSE}
penguins %>% 
  ggplot(...)
```

```{r exercise8-solution}
penguins %>%    
  ggplot(aes(x = species, y = body_mass_g)) +
  geom_boxplot(notch = TRUE)
```

### Box width

Sometimes we want the boxes to be wider or narrow than the default. You can control box width with the `width` argument.

```{r box-width, echo=TRUE, warning=FALSE}
penguins %>%    
  ggplot(aes(x = species, y = body_mass_g)) +
  geom_boxplot(notch = TRUE, width = 0.5)
```

By default, the notch width is set to be 50% of the box width. You can change that with the `notchwidth` argument.

```{r notch-width, echo=TRUE, warning=FALSE}
penguins %>%    
  ggplot(aes(x = species, y = body_mass_g)) +
  geom_boxplot(notch = TRUE, width = 0.5, notchwidth = 0.75)
```

### Wrap-up

You are now a certified distribution visualization expert!

