---
title: "Wrangling rows"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(palmerpenguins)
gradethis::gradethis_setup()
knitr::opts_chunk$set(echo = FALSE)
```

## Data wrangling

### Setup

We're working with `{dplyr}` again, so let's load the `{dplyr}` package along with the `{palmerpenguins}` package to use the `penguins` data set.

```{r echo = TRUE, message = FALSE}
library(dplyr)
library(palmerpenguins)
head(penguins)
```

Now we'll start working with rows for this lesson.


## Filtering and sorting rows

### Filtering rows {data-allow-skip=TRUE}

While base R uses the `subset()` function, tidyverse uses `filter()` to pull out rows based on data values. For instance, to filter the gentoo species of penguins:
```{r echo = TRUE}
filter(penguins, species == "Gentoo")
```

You can use multiple criteria within a column or combining multiple columns using logical operators `>`, `>=`, `<`, `<=`, `!=`, `==`, `%in%`.

**Produce a data frame of penguins with body masses between and including 4500-5000 g. **
```{r exercise1, exercise = TRUE}
filter(penguins, ...)
```
 
```{r exercise1-check}
grade_result(
  pass_if(~identical(.result, filter(penguins, body_mass_g >= 4500 & body_mass_g <=5000)))
)
```

**Produce a data frame of penguins measured in 2008 or 2009. **
```{r exercise2, exercise = TRUE}
filter(penguins, ...)
```
 
```{r exercise2-check}
grade_result(
  pass_if(~identical(.result, filter(penguins, year == 2008 | year == 2009)))
)
```

**Produce a data frame of penguins measured on Biscoe or Dream island using the `%in%` operator. **
```{r exercise3, exercise = TRUE}
filter(penguins, ...)
```
 
```{r exercise3-check}
grade_result(
  pass_if(~identical(.result, filter(penguins, island %in% c("Biscoe", "Dream"))))
)
```

**Produce a data frame of penguins of unknown sex by including only NAs. **
```{r exercise4, exercise = TRUE}
filter(penguins, ...)
```
 
```{r exercise4-check}
grade_result(
  pass_if(~identical(.result, filter(penguins, is.na(sex))))
)
```

**Produce a data frame of penguins of unknown sex by excluding those found in the vector `c("male", "female")`. Hint: This one is tricky, but remember, `!` means NOT. Feel free to check the solution if you get stuck.**
```{r exercise5, exercise = TRUE}
filter(penguins, ...)
```
 
```{r exercise5-solution}
filter(penguins, !sex %in% c("male", "female"))
```
 
```{r exercise5-check}
grade_result(
  pass_if(~identical(.result, filter(penguins, !sex %in% c("male", "female"))))
)
```


### Sorting rows

While most computations don't depend on the order of rows, some do (e.g., ranks) and sometimes it's just nice to visualize a data frame sorted by some of the column values. The `arrange()` function does this and hierarchically sorts based on the order of column names passed as arguments.

**Sort the rows by sex then body mass.**
```{r exercise6, exercise = TRUE}
arrange(penguins, ...)
```
 
```{r exercise6-check}
grade_result(
  pass_if(~identical(.result, arrange(penguins, sex, body_mass_g)))
)
```


**Now do the same, but sort body mass from highest to lowest.**
```{r exercise7, exercise = TRUE}
arrange(penguins, ...)
```
 
```{r exercise7-check}
grade_result(
  pass_if(~identical(.result, arrange(penguins, sex, desc(body_mass_g))))
)
```


## Summarizing data

### Overall summaries

While you can simply input data frame columns to functions to summarize them (e.g., `mean(penguins$bill_length_mm, na.rm = TRUE)`, you can also use `summarise()` to do this: `summarise(penguins, mean_bill_length = mean(bill_length_mm, na.rm = TRUE))`. One key advantage of using `summarise()` is that it can apply multiple functions and can apply to multiple columns.

**Calculate the mean and standard deviation of body weight (ignoring NAs) and label them as `mean` and `sd` respectively.**
```{r exercise8, exercise = TRUE}
summarise(penguins, ...)
```
 
```{r exercise8-check}
grade_result(
  pass_if(~identical(.result, summarise(penguins, mean = mean(body_mass_g, na.rm = TRUE), sd = sd(body_mass_g, na.rm = TRUE))))
)
```

**Calculate the mean of all of the bill and flipper measurements (ignoring NAs).**
```{r exercise9, exercise = TRUE}
summarise(penguins, ...)
```
 
```{r exercise9-check}
grade_result(
  pass_if(~identical(.result, summarise(penguins, across(contains("_mm"), ~mean(.x, na.rm = TRUE)))))
)
```

### Summarizing groups

The `summarise()` function is probably most useful when paired with the `group_by()` function, which allows separate summaries for each group or combination of groups. 

But since this involves applying multiple functions, before we do this, we need to review the pipe operator `%<%`. The pipe allows us to chain multiple functions together to apply them to data in the specified order. Basically, a pipe takes the output from the previous command and submits it to the next command. Therefore, we typically use `group_by` to group the data by columns then pipe the output to a `summarise()` statement to generate summaries per group.

**Calculate the mean and standard deviation of body size (labeled as `mean` and `sd` respectively) for male, female, and unknown sex penguins (ignoring NAs in body size).**
```{r exercise10, exercise = TRUE}
group_by(...) %>% 
summarise(penguins, ...)
```
 
```{r exercise10-check}
grade_result(
  pass_if(~identical(.result, group_by(penguins, sex) %>% summarise(mean = mean(body_mass_g, na.rm = TRUE), sd = sd(body_mass_g, na.rm = TRUE))))
)
```

**Calculate the mean body size (labeled as `mean`) for each species and sex (in that order) of penguin (ignoring NAs in body size).**
```{r exercise11, exercise = TRUE}
group_by(...) %>% 
summarise(penguins, ...)
```
 
```{r exercise11-check}
grade_result(
  pass_if(~identical(.result, group_by(penguins, species, sex) %>% summarise(mean = mean(body_mass_g, na.rm = TRUE))))
)
```

## Combining functions

### Using pipes

We got a taste of using the pipe with `group_by()` and `summarise()`. But it can be used to create whole chains of functions to wrangle data. The order of `filter()` and `select()` commands often doesn't matter. But where you place `summarise()` and sometimes `mutate()` and `arrange()` commands can matter. Please complete these commands in the order specified.

**Produce a data frame that includes the species, sex, island, and body mass for Dream Island penguins sorted by body mass.**
```{r exercise12, exercise = TRUE}

```
 
```{r exercise12-check}
grade_result(
  pass_if(~identical(.result, penguins %>% select(species, sex, island, body_mass_g) %>% filter(island == "Dream") %>% arrange(body_mass_g)))
)
```

**For each species, sex, and island (in that order) of penguins measured in 2007, calculate the mean flipper length (labeled `mean` and ignoring NAs).**
```{r exercise13, exercise = TRUE}

```
 
```{r exercise13-check}
grade_result(
  pass_if(~identical(.result, penguins %>% filter(year == 2007) %>% group_by(species, sex, island) %>% summarise(mean = mean(flipper_length_mm, na.rm = TRUE))))
)
```

### Wrap-up

You now have the tools for wrangling rows of data!

