---
title: "Visualizing x-y data"
output: 
  learnr::tutorial:
    progressive: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(palmerpenguins)
gradethis::gradethis_setup()
knitr::opts_chunk$set(echo = FALSE)
lincoln_weather <- read_csv("lincoln_weather.csv") %>% 
  mutate(date = as.Date(date))
lincoln_minmax_weather <- lincoln_weather %>% 
  pivot_longer(c("min_temp", "max_temp"), names_to = "measure", values_to = "temp")

```

## Introduction

This exercise covers visualizing x-y data such as associations, times series, and trends.

### Set-up

Load `{tidyverse}` and `{palmerpenguins}`.

```{r echo = TRUE, message = FALSE}
library(tidyverse)
library(palmerpenguins)
```

## Scatterplots

###

Scatterplots plot paired continuous data. Here we see (again) the penguin bill length plotted against bill depth.

```{r scatter, echo = TRUE, warning = FALSE}
penguins %>% 
  ggplot(aes(x = bill_depth_mm, y = bill_length_mm)) +
  geom_point()
```

### Adding regression lines

Just as a refresher, you can add regression lines with `geom_smooth(method = "lm")`.

```{r scatter-lm, echo = TRUE, warning = FALSE, message = FALSE}
penguins %>% 
  ggplot(aes(x = bill_depth_mm, y = bill_length_mm)) +
  geom_point() +
  geom_smooth(method = "lm")
```

###

**Now add separate regression lines for each species.**

```{r scatter-lm2, warning = FALSE, message = FALSE}
penguins %>% 
  ggplot(aes(x = bill_depth_mm, y = bill_length_mm, color = species)) +
  geom_point() +
  geom_smooth(method = "lm")
```

```{r exercise1, exercise = TRUE, warning = FALSE, message = FALSE}
penguins %>% 
  ...
```

```{r exercise1-solution}
penguins %>% 
  ggplot(aes(x = bill_depth_mm, y = bill_length_mm, color = species)) +
  geom_point() +
  geom_smooth(method = "lm")
```

### Adding diagonal lines

If you are plotting two variables that are on the same scale (e.g., pre/post, measure 1 vs. measure 2), you may want to add a diagonal line in which _x=y_. For this, we can use the `geom_abline()` function and specify the `intercept` as 0 and the `slope` as 1. Unfortunately, there are no paired variables on the same scale in the `penguins` dataset, so we'll draw a line where _2.5x=y_. That is, let's imagine that we want to plot the line showing where bill length is 2.5 times bill depth. 

**Plot the bill depth and length data including a line with slope is 2.5 and y-intercept is 0.**

```{r scatter-line, warning = FALSE}
penguins %>% 
  ggplot(aes(x = bill_depth_mm, y = bill_length_mm)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 2.5)
```

```{r exercise2, exercise = TRUE, warning = FALSE}
penguins %>% 
  ggplot(aes(x = bill_depth_mm, y = bill_length_mm)) +
  ...
```

```{r exercise2-solution}
penguins %>% 
  ggplot(aes(x = bill_depth_mm, y = bill_length_mm)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 2.5)
```

## Correlation plots

Since `{ggplot2}` does not have a built in way to easily generate a correlation plot, we're going to use base R's `pairs()` function. You can also use the `{GGally}` function `ggpairs()`.

We'll plot the correlations between all of the size measurements in `penguins`.

```{r pairs, echo = TRUE}
penguins %>% 
  select(contains("_mm") | contains("_g")) %>% 
  pairs(upper.panel = NULL)
```



## Bubble charts

### Size as a third variable

Bubble charts vary the size of points to represent information and can be used in two ways. First, they can be used with continuous data in scatterplots to represent a third variable with the size of the data points. Here we plot bill depth and length of Adelie penguins as the x and y axes. Then we map the point size to body mass to generate the bubble chart.

```{r bubble1, echo=TRUE, warning=FALSE}
penguins %>% 
  filter(species == "Adelie") %>% 
  ggplot(aes(x = bill_depth_mm, y = bill_length_mm, size = body_mass_g)) +
  geom_point()
```

###

It is difficult to see all of the individual points.

**Replicate the above plot, but make the points 50% transparent.**

```{r bubble2, warning=FALSE}
penguins %>% 
  filter(species == "Adelie") %>% 
  ggplot(aes(x = bill_depth_mm, y = bill_length_mm, size = body_mass_g)) +
  geom_point(alpha = 0.5)
```

```{r exercise3, exercise = TRUE, warning = FALSE}
penguins %>% 
  filter(species == "Adelie") %>% 
  ggplot(aes(x = bill_depth_mm, y = bill_length_mm)) +
  ...
```

```{r exercise3-solution}
penguins %>% 
  filter(species == "Adelie") %>% 
  ggplot(aes(x = bill_depth_mm, y = bill_length_mm, size = body_mass_g)) +
  geom_point(alpha = 0.5)
```

### Size as count

Bubble charts can also show the number of observations that fall within each x-y position. These can be used for both continuous and discrete data. For these plots, we use the `geom_count()` function.

```{r bubble3, echo=TRUE}
penguins %>%
  ggplot(aes(x = island, y = year)) +
  geom_count()
```

### 

In this case, year is treated as a numeric, which creates strange y-axis values. We'll learn other ways to deal with this later, but for now, let's convert year to a factor and replot.

**Coerce year to a factor and plot the above bubble chart.**

```{r bubble4}
penguins %>%
  mutate(year = factor(year)) %>% 
  ggplot(aes(x = island, y = year)) +
  geom_count()
```

```{r exercise4, exercise = TRUE, warning = FALSE}
penguins %>%
  ...
```

```{r exercise4-solution}
penguins %>%
  mutate(year = factor(year)) %>% 
  ggplot(aes(x = island, y = year)) +
  geom_count()
```

## Slopegraphs

### Plotting raw numbers

Slopegraphs are a good way to show how a continuous variable for individual participants or a categorical variable change over time or another categorical variable with few levels. Here, we'll look at how mean bill length changes as a function of year and species. First, we'll calculate the mean bill lengths, then plot the species means for each year connected by lines.

<!-- recall <- read_table(file = "http://personality-project.org/r/datasets/R.appendix3.data") -->
```{r slopegraph1, echo=TRUE, message=FALSE}
penguins %>% 
  mutate(year = factor(year)) %>% 
  group_by(species, year) %>% 
  summarise(bill_length = mean(bill_length_mm, na.rm = TRUE)) %>%
  ggplot(aes(x = year, y = bill_length, group = species)) +
  geom_point() +
  geom_line()
```

### Calculating summary statistics

Instead of calculating the means in advance, we can rely on `stat_summary()` to do the calculations in the plot.

**Reproduce the above plot using `stat_summary()` instead of `group_by()` and `summarise()` to plot means and standard errors. Remember to take the year a factor to create a reasonable x axis.**

```{r slopegraph2, warning=FALSE, message=FALSE}
penguins %>% 
  mutate(year = factor(year)) %>% 
  ggplot(aes(x = year, y = bill_length_mm, group = species)) +
  stat_summary(fun.data = mean_se) +
  stat_summary(geom = "line")
```

```{r exercise5, exercise = TRUE, warning=FALSE, message=FALSE}
penguins %>%
  ...
```

```{r exercise5-solution}
penguins %>% 
  mutate(year = factor(year)) %>% 
  ggplot(aes(x = year, y = bill_length_mm, group = species)) +
  stat_summary(fun.data = mean_se) +
  stat_summary(geom = "line")
```

###

Note that the species lines are not labelled. We're going to ignore that here and will learn how to add text labels in the adorning plots lesson.


## Time series

###

Let's explore time series data by working with the Lincoln temperature data used in _Fundamentals in Data Visualization_. First, let's import these data and convert the dates to a date format.

```{r echo=TRUE, message=FALSE}
lincoln_weather <- read_csv("lincoln_weather.csv") %>% 
  mutate(date = as.Date(date))
head(lincoln_weather)
```

We have maximum, minimum, and mean temperature for every day in 2016. We'll start by plotting daily mean temperatures as points.

```{r time-point, echo=TRUE}
lincoln_weather %>% 
  ggplot(aes(x = date, y = mean_temp)) +
  geom_point()
```

### Using lines

**Change the geometric object from points to a line with `geom_line()`.**

```{r time-line}
lincoln_weather %>% 
  ggplot(aes(x = date, y = mean_temp)) +
  geom_line()
```

```{r exercise6, exercise = TRUE, warning=FALSE, message=FALSE}
lincoln_weather %>% 
  ...
```

```{r exercise6-solution}
lincoln_weather %>% 
  ggplot(aes(x = date, y = mean_temp)) +
  geom_line()
```


### Multiple series

Let's say that we want to visualize the minimum and maximum rather than the mean. We can't do this with the current data frame because they are in different columns.

**Reshape the data frame to put the minimum and maximum (only) temperatures into a single column called `temp` and the names in a column called `measure`. Call the new data frame `lincoln_minmax_weather`.**

```{r}
lincoln_minmax_weather <- lincoln_weather %>% 
  pivot_longer(c("min_temp", "max_temp"), names_to = "measure", values_to = "temp")
```

```{r exercise7, exercise = TRUE, warning=FALSE, message=FALSE}
lincoln_minmax_weather <- 
  ...
```

```{r exercise7-solution}
lincoln_minmax_weather <- lincoln_weather %>% 
  pivot_longer(c("min_temp", "max_temp"), names_to = "measure", values_to = "temp")
```

###

Now plot the temperature as a function of date.

**Plot the time series of temperature with different colors for minimum and maximum temperature.**

```{r time-multiple}
lincoln_minmax_weather %>% 
  ggplot(aes(x = date, y = temp, color = measure)) +
  geom_line()
```

```{r exercise8, exercise = TRUE, warning=FALSE, message=FALSE}
lincoln_minmax_weather %>% 
  ...
```

```{r exercise8-solution}
lincoln_minmax_weather %>% 
  ggplot(aes(x = date, y = temp, color = measure)) +
  geom_line()
```


## Trends

###

Trends involve fitting models to capture the general patterns in data while ignoring noise. We will visualize trends with the `geom_smooth()` function, which by default uses the LOESS method of polynomial regression.

Here, we'll plot the trend in daily mean temperature.

```{r trend, echo=TRUE, message=FALSE}
lincoln_weather %>% 
  ggplot(aes(x = date, y = mean_temp)) +
  geom_point() +
  geom_smooth()
```

###

**Replicate the above plot but remove the error band by including the argument `se = FALSE`.**

```{r trend-nose, message=FALSE}
lincoln_weather %>% 
  ggplot(aes(x = date, y = mean_temp)) +
  geom_point() +
  geom_smooth(se = FALSE)
```

```{r exercise9, exercise = TRUE, warning=FALSE, message=FALSE}
lincoln_weather %>% 
  ...
```

```{r exercise9-solution}
lincoln_weather %>% 
  ggplot(aes(x = date, y = mean_temp)) +
  geom_point() +
  geom_smooth(se = FALSE)
```


### Wrap-up

You are now a certified distribution visualization expert!

