---
title: "Character strings"
output: 
  learnr::tutorial:
    progressive: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
gradethis::gradethis_setup()
knitr::opts_chunk$set(echo = FALSE)
```


## Introduction

### Character strings

First, let's load the `{stringr}` package.
```{r echo = TRUE}
library(stringr)
```


Character strings are a basic data type in R. When reading in data, typically any columns that have non-numeric data will be treated as character strings. 

As review, how do you coerce a vector of numeric data that you want to be characters? 

**Convert the vector `x` below to be a character string.**
```{r exercise1, exercise = TRUE}
x <- 1:10
x <- ...
```
 
```{r exercise1-check}
grade_result(
  pass_if(~identical(.result, x <- as.character(1:10)))
)
```

We assign strings by wrapping them in quotes. But what happens if we want to include quotes within strings?

```{r string-quotes}
question(
  'Which options would generate this character string? \n
  _The central message of Buddhism is not "every man for himself"_',
  answer('`\'The central message of Buddhism is not "every man for himself"\'`', correct = TRUE),
  answer('`\"The central message of Buddhism is not \\"every man for himself\\"\"`', correct = TRUE),
  answer('`\"The central message of Buddhism is not \"every man for himself\"\"`'),
  answer('`\'The central message of Buddhism is not \\"every man for himself\\"\'`', correct = TRUE),
  allow_retry = TRUE,
  random_answer_order = TRUE
)
```

## Working with strings

Several `{stringr}` functions allow you to work with string length, extraction, and case.

```{r string-length}
quiz(
  question(
    'What would be the output of the following command: `length(c("Avoid", "the", "green", "ones.", "They\'re", "not", "ripe", "yet."))`?',
    answer('`8`', correct = TRUE),
    answer('`5 3 5 5 7 3 4 4`'),
    answer('`"Avoid" "the" "green" "ones." "They\'re" "not" "ripe" "yet."`'),
    answer('`"Avoid the green ones. They\'re not ripe yet."`'),
    allow_retry = TRUE,
    random_answer_order = TRUE
  ),
  question(
    'What would be the output of the following command: `str_length(c("Avoid", "the", "green", "ones.", "They\'re", "not", "ripe", "yet."))`?',
    answer('`8`'),
    answer('`5 3 5 5 7 3 4 4`', correct = TRUE),
    answer('`"Avoid" "the" "green" "ones." "They\'re" "not" "ripe" "yet."`'),
    answer('`"Avoid the green ones. They\'re not ripe yet."`'),
    allow_retry = TRUE,
    random_answer_order = TRUE
  ),
  question(
    'What would be the output of the following command: `str_c(c("Avoid", "the", "green", "ones.", "They\'re", "not", "ripe", "yet."), collapse = " ")`?',
    answer('`8`'),
    answer('`5 3 5 5 7 3 4 4`'),
    answer('`"Avoid" "the" "green" "ones." "They\'re" "not" "ripe" "yet."`'),
    answer('`"Avoid the green ones. They\'re not ripe yet."`', correct = TRUE),
    allow_retry = TRUE,
    random_answer_order = TRUE
  )
)
```

**Type the code that would extract the first three characters from each element of the following vector:**
```{r exercise2, exercise = TRUE}
x <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
...
```
 
```{r exercise2-check}
grade_result(
  pass_if(~identical(.result, str_sub(x, 1, 3)))
)
```

**Type the code that converts the MLA formatted title _Beyond Good and Evil_ to APA formatted title _Beyond good and evil_.**
```{r exercise3, exercise = TRUE}
x <- "Beyond Good and Evil"
...
```
 
```{r exercise3-check}
grade_result(
  pass_if(~identical(.result, str_to_sentence(x)))
)
```

## Matching strings

`{stringr}` use regular expressions to match characters in strings. This allows us to count, find, and replace strings.

**Answer the following questions about the character string `quote <- c("Apes", "don't", "read", "philosophy")`.**
```{r string-counts}
quiz(
  question(
    'Which command would allow you to count the number of elements containing the letter `p` in  `quote`?',
    answer('`sum(str_detect(quote, "p"))`', correct = TRUE),
    answer('`quote[str_detect(quote, "p")]`'),
    answer('`str_count(quote, "p")`'),
    answer('`str_detect(quote, "p")`'),
    allow_retry = TRUE,
    random_answer_order = TRUE
  ),
  question(
    'Which command would allow you to count the number of instances of the letter `p` in each element of `quote`?',
    answer('`str_count(quote, "p")`', correct = TRUE),
    answer('`sum(str_detect(quote, "p"))`'),
    answer('`quote[str_detect(quote, "p")]`'),
    answer('`str_detect(quote, "p")`'),
    allow_retry = TRUE,
    random_answer_order = TRUE
  ),
  question(
    'Which commands would return `"Apes" "philosophy"` from `quote`?',
    answer('`str_extract(quote, "p")`'),
    answer('`str_subset(quote, "p")`', correct = TRUE),
    answer('`quote[str_detect(quote, "p")]`', correct = TRUE),
    answer('`str_detect(quote, "p")`'),
    allow_retry = TRUE,
    random_answer_order = TRUE
  ),
  question(
    'Which command would insert a `f` for each `ph` in `quote`?',
    answer('`str_replace_all(quote, "ph", "f")`', correct = TRUE),
    answer('`str_replace(quote, "ph", "f")`'),
    answer('`str_insert(quote, "ph", "f")`'),
    answer('`str_switch(quote, "ph", "f")`'),
    allow_retry = TRUE,
    random_answer_order = TRUE
  )
)
```


## Combining strings with output

### Pasting and gluing

For the following exercises, use the appropriate function to create the sentence: `"Yes they do, Otto. They just don't understand it."`

**Produce this sentence use the object `name` and the `paste()` function.**
```{r exercise4, exercise = TRUE}
name <- "Otto"
paste(...)
```
 
```{r exercise4-check}
grade_result(
  pass_if(~identical(.result, paste("Yes they do, ", name, ". They just don't understand it.", sep = "")))
)
```

**Now produce the sentence using the `paste0()` function.**
```{r exercise5, exercise = TRUE}
name <- "Otto"
paste(...)
```
 
```{r exercise5-check}
grade_result(
  pass_if(~identical(.result, paste0("Yes they do, ", name, ". They just don't understand it.")))
)
```

**Now produce the sentence using the `str_glue()` function with `name` as an argument.**
```{r exercise6, exercise = TRUE}
str_glue(...)
```
 
```{r exercise6-check}
grade_result(
  pass_if(~identical(.result, str_glue("Yes they do, {name}. They just don't understand it.", name = "Otto")))
)
```

### Wrap-up

You're now an expert stringer!

