---
title: "Importing and exporting data"
output: 
  learnr::tutorial:
    progressive: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
```


## Importing data

There are numerous ways of importing data depending on if you have the files locally on your computer or if they are available on the internet. Different functions (and packages) are needed to import data based on it's location and file type. The two functions you will probably use most often are base R's `read.csv()` and `{readr}`'s `read_csv()`. I recommend reading the documentation for both of these (`?read.csv()` and `?readr::read_csv`).

And don't forget that `here::here()` will help facilitate data import when you are using R Projects (which you should!). 

```{r importing-quiz}
quiz(
question(
  'Which of the following are plain text files?',
  answer('`.txt`', correct = TRUE),
  answer('`.csv`', correct = TRUE),
  answer('`.R`', correct = TRUE),
  answer('`.Rmd`', correct = TRUE),
  answer('`.rds`'),
  answer('`.Rdata`'),
  answer('`.xlsx`'),
  answer('`.sav`'),
  allow_retry = TRUE,
  random_answer_order = TRUE
),
question(
  'Which of the following would produce a tibble?',
  answer('`read_csv()`', correct = TRUE),
  answer('`read.csv()`', message = "Oops, that produces a data frame."),
  answer('`write.csv()`', message = "Oops, that writes a data frame to a CSV file."),
  answer('`write_csv()`', message = "Oops, that writes a data frame to a CSV file."),
  allow_retry = TRUE,
  random_answer_order = TRUE
),
question_text(
  "Write the code that _imports_ a fictional `mydata.csv` file from the path `data/raw_data/` using base R syntax but does not use the `{here}` package.",
  answer('read.csv(file = "data/raw_data/mydata.csv")', correct = TRUE),
  answer('read.csv("data/raw_data/mydata.csv")', correct = TRUE),
  answer('read.csv(file = "data/raw_data/mydata")', message = "Oops, you forgot the `.csv` extension."),
  answer('read.csv("data/raw_data/mydata")', message = "Oops, you forgot the `.csv` extension."),
  answer('read.csv(file = "mydata.csv")', message = "Oops, you forgot the path."),
  answer('read.csv("mydata.csv")', message = "Oops, you forgot the path."),
  answer('read.csv(file = here("data/raw_data/mydata.csv"))', message = "Oops, that uses `{here}`."),
  answer('read.csv(here("data/raw_data/mydata.csv"))', message = "Oops, that uses `{here}`."),
  answer('read_csv(file = "data/raw_data/mydata.csv")', message = "Oops, that uses `{readr}'s `read_csv()` function."),
  answer('read_csv("data/raw_data/mydata.csv")', message = "Oops, that uses `{readr}'s `read_csv()` function."),
  allow_retry = TRUE,
  random_answer_order = TRUE
)
)
```


## Exporting data

Typically, exporting data to files is more straightforward because most of the time, you will be exporting CSV files. Again, the two functions you will probably use most often are base R's `write.csv()` and `{readr}`'s `write_csv()` (`?write.csv()` and `?readr::write.csv`).

```{r exporting-quiz}
quiz(
question(
  'Which of the following is a default argument value for `write.csv()`?',
  answer('`row.names = TRUE`', correct = TRUE),
  answer('`sep = ";"`'),
  answer('`dec = ","`'),
  answer('`quote = FALSE`'),
  allow_retry = TRUE,
  random_answer_order = TRUE
),
question(
  'Which of the following would save your workspace?',
  answer('`save.image("workspace.Rdata")`', correct = TRUE),
  answer('`save("workspace.Rdata")`', message = "Oops, that uses the wrong function."),
  answer('`save.image("workspace.rds")`', message = "Oops, that saves to the wrong file type."),
  answer('`saveRDS("workspace.rds")`', message = "Oops, that uses the wrong function."),
  allow_retry = TRUE,
  random_answer_order = TRUE
),
question_text(
  "Write the code that _exports_ the `mydata` object to a file called `mydata2.csv` in the path `data/` using `{readr}` syntax and using the `here::here()` function.",
  answer('write_csv(mydata, file = here("data/mydata2.csv"))', correct = TRUE),
  answer('write_csv(mydata, here("data/mydata2.csv"))', correct = TRUE),
  answer('write_csv(mydata, file = here::here("data/mydata2.csv"))', correct = TRUE),
  answer('write_csv(mydata, here::here("data/mydata2.csv"))', correct = TRUE),
  answer('write_csv(mydata, file = "data/mydata2.csv")', message = "Oops, you forget to use `here()`."),
  answer('write_csv(mydata, "data/mydata2.csv")', message = "Oops, you forget to use `here()`."),
  allow_retry = TRUE,
  random_answer_order = TRUE
)
)
```

## Data validation

Validating that your data are as you expect is a critical step before processing them. Minimally, you should inspect your data sets with `str()` or `glimpse()` or look at them in the _Environment_ tab. You can use the `{validate}` package to build validation rules to check your data. You can also use more elaborate testing with assertions and unit testing.

### Validating with `{dataReporter}`

The [`{dataReporter}`](https://github.com/ekstroem/dataReporter) package provides a nice way of generating a data dictionary (they call them codebooks) while giving you a nice overview of your data. Install and load [`{dataReporter}`](https://github.com/ekstroem/dataReporter). Run `makeCodebook()` on `toyData` to explore this data set.


```{r validation-quiz}
quiz(
question(
  'How many rows does the `toyData` have?',
  answer('15', correct = TRUE),
  answer('6'),
  answer('7'),
  answer('9'),
  allow_retry = TRUE,
  random_answer_order = TRUE
),
question(
  'What data type is the `pill` column?',
  answer('factor', correct = TRUE),
  answer('numeric'),
  answer('character'),
  answer('logical'),
  allow_retry = TRUE,
  random_answer_order = TRUE
),
question(
  'How many unique values does the `events` column have?',
  answer('9', correct = TRUE),
  answer('15'),
  answer('7'),
  answer('1'),
  allow_retry = TRUE,
  random_answer_order = TRUE
),
question(
  'What percent of the `events` column is missing?',
  answer('20%', correct = TRUE),
  answer('13.33%'),
  answer('0%'),
  answer('100%'),
  allow_retry = TRUE,
  random_answer_order = TRUE
),
question(
  'What is the median value of the `change` column?',
  answer('0.33', correct = TRUE),
  answer('-0.62'),
  answer('0.66'),
  answer('1.6'),
  allow_retry = TRUE,
  random_answer_order = TRUE
),
question(
  'Which variable is described as a key with distinct values for each observation?',
  answer('`id`', correct = TRUE),
  answer('`region`'),
  answer('`events`'),
  answer('`pill`'),
  answer('`spotifysong`'),
  allow_retry = TRUE,
  random_answer_order = TRUE
)
)
```


### Wrap-up

Nice---you should be able to import and validate your data now.

