---
title: "Tutorial"
output: 
  learnr::tutorial:
    progressive: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(palmerpenguins)
library(lubridate)
gradethis::gradethis_setup()
knitr::opts_chunk$set(echo = FALSE)
```


## Introduction

### Factors and dates

As a reminder, the four core data types are character, double, integer, and logical. Factors and dates are _augmented vectors_. Factors are augmented integers and dates are augmented doubles.

## Factors

### Set-up
To work with factors, we'll use the `{forcats}` package, which is automatically loaded when you load the core tidyverse packages. We'll go ahead and load `{tidyverse}` since we'll need other tidyverse packages. We'll also load the `{palmerpenguins}` package to work with those data.

```{r echo = TRUE}
library(tidyverse)
library(palmerpenguins)
```

### Ordering factor levels
The `penguins` data frame has three variables that are factors: species, island, and sex.

```{r}
penguins
```

First, let's check the levels for the `species` variable.

```{r echo = TRUE}
levels(penguins$species)
```

They seem to be in alphabetical order. Let's say that we want the levels to be in the opposite order: Gentoo, Chinstrap, Adelie.

**Reverse the order of the levels by replacing the `...` with the proper `{forcats}` function. Note that since we are just working with the vector `penguins$species`, we don't need to mutate anything---we can just use one of the `{forcats}` functions to change the levels.**
```{r exercise1, exercise = TRUE}
penguins$species %>% 
  ... %>% 
  levels()
```
 
```{r exercise1-solution}
penguins$species %>% 
  fct_rev() %>% 
  levels()
```

```{r exercise1-check}
grade_result(
  pass_if(~identical(.result, penguins$species %>% 
  fct_rev() %>% 
  levels()))
)
```

**Now put the levels in the order Chinstrap, Gentoo, Adelie.**
```{r exercise1b, exercise = TRUE}
penguins$species %>% 
  ... %>% 
  levels()
```
 
```{r exercise1b-solution}
penguins$species %>% 
  fct_relevel("Chinstrap", "Gentoo", "Adelie") %>% 
  levels()
```

```{r exercise1b-check}
grade_result(
  pass_if(~identical(.result, penguins$species %>% 
  fct_relevel("Chinstrap", "Gentoo", "Adelie") %>% 
  levels()
  ))
)
```

### Recoding factors

Now we want the word "Island" to be added to each of the island names in the data frame. So we need to recode the existing factor levels.

**Recode the island variable to add "Island" to each factor level (e.g., "Biscoe Island", "Dream Island", and "Torgersen Island") using a `{forcats}` function (not a `{stringr}` function).**
```{r exercise2, exercise = TRUE}
penguins$island %>% 
  ... %>% 
  levels()
```
 
```{r exercise2-solution}
penguins$island %>% 
  fct_recode("Biscoe Island" = "Biscoe", "Dream Island" = "Dream", "Torgersen Island" = "Torgersen") %>% 
  levels()
```

```{r exercise2-check}
grade_result(
  pass_if(~identical(.result, penguins$island %>% 
  fct_recode("Biscoe Island" = "Biscoe", "Dream Island" = "Dream", "Torgersen Island" = "Torgersen") %>% 
  levels()))
)
```

###

Now we want to create a new column that provides each species' diet. Gentoo penguins are generalists (meaning they feed on a number of different food items) and Adelie and chinstrap penguins are specialists that feed primarily on krill. To do this, we will mutate the `species` variable into a new `diet` variable that recodes Gentoo to "generalist" and Adelie and chinstrap to "specialist".

**Create a new column that recodes the species to their diets using the `fct_collapse()` function.**
```{r exercise3, exercise = TRUE}
penguins %>% 
  mutate(diet = fct_collapse(...))
```
 
```{r exercise3-solution}
penguins %>% 
  mutate(diet = fct_collapse(species, generalist = "Gentoo", specialist = c("Adelie", "Chinstrap")))
```

```{r exercise3-check}
grade_result(
  pass_if(~identical(.result, penguins %>% 
  mutate(diet = fct_collapse(species, generalist = "Gentoo", specialist = c("Adelie", "Chinstrap")))))
)
```

## Dates

### Set-up

Working with dates requires the `{lubridate}` package, which is not part of the core tidyverse packages. Therefore, we need to load it separately.

```{r echo = TRUE, message = FALSE}
library(lubridate)
```

### Today's date

```{r todays-date}
question(
  'Which of the following commands return today\'s date (and only the date)?',
  answer('`Sys.Date()`', correct = TRUE),
  answer('`today()`', correct = TRUE),
  answer('`Sys.Time()`'),
  answer('`Sys.date()`'),
  answer('`Sys.time()`'),
  answer('`now()`'),
  allow_retry = TRUE,
  random_answer_order = TRUE
)
```

### Creating dates and times

Remember the standard way to format dates is the ISO 8601 format: YYYY-MM-DD. Type the code to convert the following dates or date/times to the standard format (e.g., 2021-06-11).

**11 Jun 21**
```{r exercise4, exercise = TRUE}
mydate <- "11 Jun 21"
...
```
 
```{r exercise4-check}
grade_result(
  pass_if(~identical(.result, dmy(mydate)))
)
```

**2021-6-11**
```{r exercise5, exercise = TRUE}
mydate <- "2021-6-11"
...
```
 
```{r exercise5-check}
grade_result(
  pass_if(~identical(.result, ymd(mydate)))
)
```

**June 11th 2021**
```{r exercise6, exercise = TRUE}
mydate <- "June 11th 2021"
...
```
 
```{r exercise6-check}
grade_result(
  pass_if(~identical(.result, mdy(mydate)))
)
```

**06/11/2021 15:00**
```{r exercise7, exercise = TRUE}
mydate <- "06/11/2021 15:00"
...
```
 
```{r exercise7-check}
grade_result(
  pass_if(~identical(.result, mdy_hm(mydate)))
)
```

### Extracting date/time components

Once your dates and times are formatted correctly, it is easy to extract individual components with the provided functions in `{lubridate}`.

**Find today's date and extract the full word for the month (no abbreviations) using a `{lubridate}` function.**
```{r exercise8, exercise = TRUE}
mydate <- today()
...
```
 
```{r exercise8-check}
grade_result(
  pass_if(~identical(.result, month(mydate, label = TRUE, abbr = FALSE)))
)
```

**Find today's date and extract the abbreviation for the day of the week using a `{lubridate}` function.**
```{r exercise9, exercise = TRUE}
mydate <- today()
...
```
 
```{r exercise9-check}
grade_result(
  pass_if(~identical(.result, wday(mydate, label = TRUE)))
)
```

**Create a character vector (not factors) of the full names of the _weekdays_ (not weekend days) using a `{lubridate}` function.**
```{r exercise10, exercise = TRUE}
...
```
 
```{r exercise10-solution}
as.character(wday(2:6, label = TRUE, abbr = FALSE))
```

```{r exercise10-check}
grade_result(
  pass_if(~identical(.result, as.character(wday(2:6, label = TRUE, abbr = FALSE))))
)
```

### Time spans

One of the great things about working with dates is that you can use math to calculate time spans.

**Calculate the number of days that have passed since the class started on May 17th, 2021.**
```{r exercise11, exercise = TRUE}
...
```
 
```{r exercise11-check}
grade_result(
  pass_if(~identical(.result, today() - ymd("2021-05-17")))
)
```

**Find the date 123 days from now.**
```{r exercise12, exercise = TRUE}
...
```
 
```{r exercise12-check}
grade_result(
  pass_if(~identical(.result, today() + 123))
)
```

### Wrap-up

You know your factors and dates!

