---
title: "Course packages"
author: "Jeffrey R. Stevens"
date: "`r Sys.Date()`"
urlcolor: blue
output:
  pdf_document: 
    includes:
      in_header: columns.tex

---

:::::: {.cols data-latex=""}

::: {.col data-latex="{0.55\textwidth}"}

## Getting started
* [{devtools}](https://devtools.r-lib.org)
* [{here}](https://here.r-lib.org/)
* [{tidyverse}](https://www.tidyverse.org/)

## Coding
* [{lintr}](https://github.com/jimhester/lintr)
* [{palmerpenguins}](https://allisonhorst.github.io/palmerpenguins/)
* [{styler}](https://styler.r-lib.org/)

## Version control
* [{gitcreds}](https://r-lib.github.io/gitcreds/)

## Literate programming
* [{knitr}](https://yihui.org/knitr/)
* [{papaja}](https://github.com/crsh/papaja)
* [{rmarkdown}](https://rmarkdown.rstudio.com/docs/)
* [{tinytex}](https://yihui.org/tinytex/)

:::

::: {.col data-latex="{0.05\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::

::: {.col data-latex="{0.4\textwidth}"}

## Importing and exporting data
* [{dataReporter}](https://github.com/ekstroem/dataReporter)
* [{qualtRics}](https://docs.ropensci.org/qualtRics/)
* [{readxl}](https://readxl.tidyverse.org/)

## Data wrangling
* [{nycflights13}](https://github.com/tidyverse/nycflights13)

## Dates
* [{lubridate}](https://lubridate.tidyverse.org/)

## Plotting data
* [{cowplot}](https://wilkelab.org/cowplot/)
* [{ggthemes}](https://github.com/jrnold/ggthemes)
* [{patchwork}](https://patchwork.data-imaginist.com/)

## Tables
* [{gt}](https://gt.rstudio.com/)

:::
::::::

# Installing packages

If you are using Windows, install [RTools](https://cran.r-project.org/bin/windows/Rtools/) before proceeding.

Then copy and paste the following line to install all (hopefully!) packages needed in the course. **Warning: it may take a while for these to install, so don't start this if you need access to your R session (or open another R session to do this).**


```{r packages, tidy=TRUE, tidy.opts=list(width.cutoff=80), eval = FALSE}
install.packages(c("tidyverse", "devtools", "here", "lintr", "palmerpenguins", "styler", "gitcreds", "tinytex", "knitr", "rmarkdown", "dataReporter", "qualtRics", "readxl", "nycflights13", "lubridate", "cowplot", "ggthemes", "patchwork", "gt"))
```

If you experience errors installing any, try installing them individually.


Installing {papaja} requires (1) already having {tinytex} installed and (2) using `install_github()`:

```{r papaja, eval = FALSE}
devtools::install_github("crsh/papaja")
```
